Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-09-17 Final_Manuscript.Rmd
finalizing article

# Results
## Data analysis
```{r create_r-references, warning=FALSE, message=FALSE}
r_refs(file = "r-references.bib")

@@

## Analysis Strategy
```{r apim, include=TRUE, echo=FALSE, results='asis', out.width="400px", fig.cap="Basic actor-partner interdependence model (APIM) depiction."}
include_graphics("APIM_figure.png")
```

@@

##Main Results
```{r semfigure, include=TRUE, echo=FALSE, results='asis', out.width="400px", fig.cap="Path Analysis Model with Estimates"}
include_graphics("SEMfigure.png")
```

@@

#Discussion
The current study tested whether the model of interpersonal objectification and state self-objectification (SSO) used in @garcia2016objectification replicates in a same-gender dyadic sample of women only. Although past research has found that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], this is the first study to test if *interpersonal* other-objectification by women is related to state self-objectification in their woman-identified interaction partners. As hypothesized, the current study did find a significant relationship between a woman's interaction partner's report of having objectified her and her own post-interaction feelings of self-objectification. That is, there was a significant partner effect of other-objectification on SSO. This effect extends the equivalent relationship found in mixed-gender interactions to the context of same-gender interactions between women. Thus, evidence suggests that it is not only the male gaze, and male other-objectification that is related to women's state self-objectification, but being objectified by another woman can also result in SSO, at least in the context of a scenario where they know they are being evaluated as a potential dating partner.

MORE ABOUT THIS PARTNER EFFECT

As is the case in all correlational studies, we cannot be sure about the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation is theoretically interesting given that, in both the @garcia2016objectification study and in the current study, it is *women's* state self-objectification that relates to being objectified by one's partner. However, Objectification Theory [@robertsfredrickson], as well as past experimental studies [@saguyetal2010], suggest that the causal flow is from other-objectification to SSO.

Where the results of the current study diverge most notably from the results of studies testing interpersonal objectification among mixed-gender dyads is the lack of evidence for relationships between SSO and interaction inauthenticity. In addition, we found no evidence of *direct* or *indirect* relationships between SSO and cognitive functioning, relationship agency, and career aspirations. Nor was there evidence of *direct* or *indirect* relationships between other-objectification and any of these outcomes. This is somewhat suprising given the extant evidence linking SSO and cognitive functioning [@fredrickson1998swimsuit; @quinn2011performance; and see @moradi2008 for a review] and the research on interpersonal other-ojectification and cognitive functioning [@logel2009interacting; @garcia2016objectification; @gervais2011you]. This lack of evidence could potentially signal diverging processes between women's experiences with interpersonal objectification from men and interpersonal objectification from women. There is quite a bit of evidence suggesting that the male gaze is particularly detrimental [@calogero2004test; @robertsfredrickson; @gervais2011you; @gervais2013my], and perhaps the self-objectification experienced within an interaction with a women is not as harmful as the self-objectification experienced within an interaction with a man. However, as a strong note of caution, it is best not to interpret a null result as evidence of no relationship.

The lack of evidence for a relationship between SSO and interaction authenticity is surprising and again, should not be interpreted as evidence of no relationship. It should be noted that the estimate of this relationship was very small, but in the negative direction, as anticipated. If it is the case that there is a smaller (i.e., weaker) connection between women's feelings of SSO and inauthenticity in interactions with other women than in interactions with men, models of interpersonal objectification, like the SIMO [@gervais2020social], could be extended by including inauthenticity as a potential mediating factor. Further, gender of the objectifyer/interaction partner could also be added to the SIMO model to help understand the circumstances that require other-objectification and SSO to have negative consequences for women.

Although the urrent study did not find a connection between SSO and authenticity, we did find significant positive relationships between authenticity and relationship agency, and authenticity and career aspirations. The relationship between authenticity and cognitive functioning was also estimated as positive, but was not statistically signficant. Again, due to the lack of connection between SSO and authenticity, we found no evidence of *indirect* relationships between SSO and these outcome variables. This evidence of a relationship between authenticity and the relational outcome variables (i.e., relationship agency and career aspirations) provides evidence that corroborates past findings that felt authenticity in interactions is important for healthy relationship functioning [@garcia2016objectification; @teran2020relational]. Just as authenticity has been found to be important in intergroup interactions [@brunelletal2010; @garcia2016objectification; @tolman2006looking], we again find more evidence here that disruptions in feelings of authenticity can negatively impact relationships beyond the current partner. Although the current study did not find a connection between authenticity and SSO, this seems theoretically to be a natural connection, and future work must explore when and how SSO leads to inauthenticity in interpersonal objectification situations.

@@

###Sample Characteristics
In addition to being small and combined across two institutions, the sample of the current study was comprised of Western women, being that sexual objectification is most prevalent in this culture [@loughnan2015], and research on objectification conducted outside of Western or Westernized countries is scarce [@moradi2008], although more current work examines objectification from a cross-cultural framework [@loughnan2015, @wollast2020cultural]. Since “bodies exist within social and cultural contexts, and hence are also constructed through sociocultural practices and discourses” [@robertsfredrickson, p.174], it is important to consider how diverse social identities within unique cultural contexts may inform sexual objectification phenomenon to test the cross-cultural applicability of theoretical frameworks [@loughnan2015]. Further, sexualizing experiences and self-objectification are thought to begin a very young age, and thus, researchers have only recently begun to examine such experiences among children [e.g., @holland2016; @jongenelis2014; @bury2016]. Considering the fact that the average mean age of the investigated participants of this current study was 18.85 years, research among younger and older individuals is needed, especially because self-objectification may change over time [@robertsfredrickson]. It may be valuable to question the extent to which children, adolescents, or emerging adults of different races or ethnicities are exposed to varied amounts of sexualizing content. [Kat: Or a longitudinal study is needed.]

Men-men

###Sexual Objectification and the Objectifyer's Gender
The current sample contained a mixture of heterosexual and non-heterosexual women, but all participants were asked to think about and evaluate their partner as a potential dating/romantic partner. We think that hetereosexual women are able to do this with other women as their target---and there is evidence that they might do this readily [@strelan2005women; @puvia2013being]---they may be even more apt to activate social comparison processes [@festinger1954theory] than women who are sexually attracted to other women (non-heterosexual women). This differential psychological process between women with differing sexualities might have served to dampen our ability to detect relationships, adding variability. [Kat: The only other thing I would add re: sexuality limitations are that we didn’t collect/use sexual orientation, so we couldn’t analyze results by that at all/our samples might have differed there.]

@@

###Measurement
Future experiments or longitudinal studies should explore the external validity of the notions of self-objectification and how the operationalization of self-objectification may be improved.

###Alternative Model
@puvia2013being alternative model. TSO -> SOO, mediated by SSO. “Found that women’s motivation to look attractive (for men) and internalize sociocultural beauty standards were positively linked with the dehumanization of sexually objectified female targets.” Tis relationship was mediated by their own level of self-objectification.”

##Clinical Implications
The results from the current analysis highlight how subtle forms of sexist discrimination operate, and provide valuable insight into prevention and intervention efforts in both clinical and educational contexts. These results are quite useful for promoting mental health, the creation and maintence of early action programs for girls and young women, and for scholars and practitioners to work intentially to provide the tools necessary to circumvent or mitigate negative effects on self-objectification to combat such experiences.

#Conclusion
The results of this study demonstrate the complex and ambivalent nature of sexual objectification and additionally highlight the psychological and social consequences of such objectification processes on women’s social relationships and well-being. We did not find a significant effect between actor SSO and felt authenticity in the interaction, which suggests that there is not sufficient evidence to support the claim that partner objectification is the cause for the diverse range of negative effects related to interaction inauthenticity.

>>>>

# Results
## Analysis Strategy
```{r create_r-references, warning=FALSE, message=FALSE}
r_refs(file = "r-references.bib")

@@

```{r apim, include=TRUE, echo=FALSE, results='asis', out.width="400px", fig.cap="Basic actor-partner interdependence model (APIM) depiction."}
include_graphics("APIM_figure.png")
```

@@

## Main Results
```{r semfigure, include=TRUE, echo=FALSE, results='asis', out.width="400px", fig.cap="Path Analysis Model with Estimates"}
include_graphics("SEMfigure.png")
```

@@

# Discussion
The current study tested whether the model of interpersonal objectification and state self-objectification (SSO) used in @garcia2016objectification replicates in a same-gender dyadic sample of women. Although past research has found that women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], this is the first study to test if *interpersonal* other-objectification by women is related to state self-objectification in their woman-identified interaction partners. As hypothesized, the current study did find a significant relationship between a woman's interaction partner's report of having objectified her and her own post-interaction feelings of self-objectification. That is, there was a significant partner effect of other-objectification on SSO. This effect extends the equivalent relationship found in mixed-gender interactions to the context of same-gender interactions between women. Thus, evidence suggests that it is not only the male gaze, and male other-objectification that is related to women's state self-objectification, but being objectified by another woman is also related to women's SSO, at least in the context of a scenario where they know they are being evaluated as a potential dating partner.

As is the case in all correlational studies, we cannot be sure about the causal direction. The model used by the current study asserts that being objectified by a woman causes women to self-objectify, but it could be that women's SSO causes them to be objectified by their interaction partner. This latter interpretation is theoretically interesting given that, in both the @garcia2016objectification study and in the current study, it is *women's* state self-objectification that relates to being objectified by one's partner. Objectification Theory [@robertsfredrickson], as well as some past experimental studies [@saguyetal2010] suggest that at least one direction of the causal flow is from other-objectification to SSO. Other studies, especailly those investigating explicitly women objectifying other women [@puvia2013being], have relately found evidence that women's *trait* self-objectification causes women to objectify (dehumanize) other women and this link is mediated by *state* self-objectification. It could follow that we tend to objectify other women who objectify us and through a process of TSO causes SSO which in turn causes objectifation of one's partner, which in turn causes one's partner to objectify us. The current study does not provide any evidence of this alternative model, however---TSO was not significantly correlated with SSO and actor's objectification was not correlated with partner's objectification.

Where the results of the current study diverge most notably from the results of studies testing interpersonal objectification among mixed-gender dyads is the lack of evidence for relationships between SSO and interaction inauthenticity. In addition, we found no evidence of *direct* or *indirect* relationships between SSO and cognitive functioning, relationship agency, and career aspirations. Nor was there evidence of *direct* or *indirect* relationships between other-objectification and any of these outcomes. This is somewhat suprising given the extant evidence linking SSO and cognitive functioning [@fredrickson1998swimsuit; @quinn2011performance; and see @moradi2008 for a review] and the research on interpersonal other-ojectification and cognitive functioning [@logel2009interacting; @garcia2016objectification; @gervais2011you]. This lack of evidence could potentially signal diverging processes between women's experiences with interpersonal objectification from men and interpersonal objectification from women. There is quite a bit of evidence suggesting that the male gaze is particularly detrimental [@calogero2004test; @robertsfredrickson; @gervais2011you; @gervais2013my], and perhaps the self-objectification experienced within an interaction with a women is qualtatively different, and perhaps not as harmful, as the self-objectification experienced within an interaction with a man. However, as a strong note of caution, it is best not to interpret a null result as evidence of no relationship and more research on interpersonal objectification among women is needed.

The lack of evidence for a relationship between SSO and interaction authenticity is surprising and again, should not be interpreted as evidence of no relationship. It should be noted that the estimate of this relationship was small (close to zero), but in the negative direction. If it is the case that there is a smaller (i.e., weaker) connection between women's feelings of SSO and inauthenticity in interactions with other women than in interactions with men, models of interpersonal objectification, like the SIMO [@gervais2020social], could be extended by including inauthenticity as a potential mediating factor. Further, gender of the objectifyer/interaction partner could be added as a moderator of the self-objectificaiton to authenticity process to help understand the circumstances that require other-objectification and SSO to have negative consequences for women.

Although the current study did not find a connection between SSO and authenticity, we did find significant positive relationships between authenticity and relationship agency, and authenticity and career aspirations. The relationship between authenticity and cognitive functioning was also estimated as positive, but was not statistically signficant. Again, due to the lack of connection between SSO and authenticity, we found no evidence of *indirect* relationships between SSO and these outcome variables. This evidence of a relationship between authenticity and the relational outcome variables (i.e., relationship agency and career aspirations) provides evidence that corroborates past findings that felt authenticity in interactions is important for healthy relationship functioning [@garcia2016objectification; @teran2020relational]. Just as authenticity has been found to be important in intergroup interactions [@brunelletal2010; @garcia2016objectification; @tolman2006looking], we again find more evidence here that disruptions in feelings of authenticity can negatively impact relationships beyond the current partner. Although the current study did not find a connection between authenticity and SSO, this seems theoretically to be a natural connection, and future work must explore when and how SSO leads to inauthenticity in interpersonal objectification situations.

@@

###Sample Characteristics
In addition to being relatively small and combined across two institutions, another limitation of the current study is that the sample was comprised of Western women only. Being that self-objectification is most prevalent in Western culture [@loughnan2015] research on objectification conducted outside of Western or Westernized countries has been scarce [@moradi2008], although more recent work has examined objectification from a cross-cultural framework [@loughnan2015, @wollast2020cultural]. Since “bodies exist within social and cultural contexts, and hence are also constructed through sociocultural practices and discourses” [@robertsfredrickson, p.174], it is important to consider how diverse social identities within unique cultural contexts may inform sexual objectification phenomenon to test the cross-cultural applicability of theoretical frameworks [@loughnan2015].

Further, sexualizing experiences and self-objectification are thought to begin a very young age, and thus, researchers have only recently begun to examine such experiences among children [e.g., @holland2016; @jongenelis2014; @bury2016]. Considering the fact that the average mean age of the investigated participants of this current study was 18.85 years, research among younger and older individuals is needed, especially because self-objectification may change over time [@robertsfredrickson]. It may be valuable to question the extent to which children, adolescents, or emerging adults of different races or ethnicities are exposed to varied amounts of sexualizing content. Further, more longitudinal studies investigating the developmental experience of interpersonal objectification and self-objectificaion are needed [see CITE and CITE for exmaples].

###Sexual Objectification and the Objectifyer's Gender
The current sample contained a mixture of heterosexual and non-heterosexual women, but all participants were asked to think about and evaluate their partner as a potential dating/romantic partner. We think that hetereosexual women are able to do this with other women as their target---and there is evidence that they might do this readily [@strelan2005women; @puvia2013being]---they may be even more apt to activate social comparison processes [@festinger1954theory] than women who are sexually attracted to other women (non-heterosexual women). This differential psychological process between women with differing sexualities might have served to dampen our ability to detect relationships, adding variability.

@@

#Conclusion
The results from the current analysis highlight how subtle forms of sexist discrimination operate, and provide valuable insight into prevention and intervention efforts in both clinical and educational contexts. These results are quite useful for promoting mental health, the creation and maintence of early action programs for girls and young women, and for scholars and practitioners to work intentially to provide the tools necessary to circumvent or mitigate negative effects on self-objectification to combat such experiences.

The results of this study demonstrate the complex and ambivalent nature of sexual objectification and additionally highlight the psychological and social consequences of such objectification processes on women’s social relationships and well-being. We did not find a significant effect between actor SSO and felt authenticity in the interaction, which suggests that there is not sufficient evidence to support the claim that partner objectification is the cause for the diverse range of negative effects related to interaction inauthenticity.
